---
title: "Project Two"
output:
  pdf_document: default
  html_document: default
---

Due Oct. 13 at 11:59 PM. 

download.file(url = "https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2023/main/vignettes/project_two.Rmd", destfile = "/cloud/project/projects/project_two.Rmd")

For this first part of the exam, you can either use `surveys_complete.csv` or your own data. If you are using your own data, you must have data in which you think you have a numerical predictor variable and a numerical response variable. If you are using `surveys_complete`, you can use weight and hindfoot_length for this.

Save this file in your `projects` directory. You can either save it in a project two subdirectory, or just put it in projects. Either way is fine.


1) Load in your data. Which variable will you be using as a predictor, and which as a response? (5 pts)

```{r}
surveys_complete <- surveys %>% 
  na.omit()

surveys_complete
```

```
Hindfoot_length is the predictor, and weight is the response. 
```

2) Plot the two against each other with a scatter plot. Do the data appear to be related linearly? (5 pts)


```{r}
ggplot(data = surveys_complete, mapping = aes(x = hindfoot_length, y = weight)) + geom_point()
```

```
No, the data does not appear to be related linearly.
```


3) Fit the linear model. View the summary. (5 pts)


```{r}
model_fit <- lm(weight ~ hindfoot_length, data = surveys_complete) 
summary(model_fit)
```

4) Does the summary make sense when you compare it to the plot you made? Does our model have good predictive power? Evaluate the residual standard error, intercept, and R-Squared in particular. Would you say your predictor predicts the response?  (10 pts)


```
The summary does make sense when it is compared to the plot that was made. I would say that the model does not have good predictive power. When looking at the summary, the residual standard error is high, being that it is 0.47884 for the intercept and 0.01558 for the hindfoot_length. When hindfoot_length is 0, the intercept (weight) is -32.87047. This number is not possible for weight. Lastly, the R-squared is 0.4673, when it should be much closer to 1 in order to be a good predictive power. Because of this, our predictor does not predict the response effectively.
```


5) Plot the model on the graph. Increase the size of the text so it is comfortably readable at 5 feet. Make sure axis labels are readable and not overlapping with one another. (5 pts)

```
ggplot(data = surveys_complete, mapping = aes(x = hindfoot_length, y = weight)) + geom_point() + theme(text = element_text(size = 30))
```


6) Check the normality of the residuals. Do they look OK? Are we violating assumptions? (5 pts)

```{r}

augmented_fit <- broom::augment(model_fit)
qqnorm(augmented_fit$.resid)

The normality of the residuals do not look okay, and we are violating assumptions, because we have a curved line. If the residuals were normal and assumptions were not being violated, the line would be straight. 

```

Why is normality of residuals important? 

```{r}
The normality of the residuals is important, because it determines how accurate hindfoot_length predicts weight.
```

7) If you are using `surveys_complete`: Is there interspecific variation in the linear model? Hint: look at our prior work with facet plots. (15 pts) 

If you are *not* using  `surveys_complete`: Do you think there are groupings within your data that may have a separate linear model? Try grouping the data by that variable and redoing the lm. If this would not make sense for your data, you may also attempt to do multiple predictor variables. (15 pts)

```{r}
ggplot(data = surveys_complete, mapping = aes(x = hindfoot_length, y = weight)) + geom_col() + facet_wrap(facets = vars(species_id)) +theme_light()

There is interspecific variation in the linear model. 
```

## Part Two

In this portion, you are free to use your own data if you have a categorical predictor variable and a response variable. Otherwise, you may use the columns sex and weight in `surveys_complete`

1) First, plot the data grouped by sex (5 pts)

```{r}
ggplot(data = surveys_complete, mapping = aes(x = sex, y = weight, color = sex)) + geom_point()
```

2) Try an ANOVA of this data (5 pt)

```{r}
model_fit <- lm(weight ~ sex, data = surveys_complete)
aov(model_fit) -> anova_model_fit
summary(anova_model_fit)
```

3) How about a linear model? What information does this give you that an ANOVA can't? (5 pts)


```{r}
summary(model_fit)

```

```
Linear model gives you R-squared values, significant codes, and coefficients that the ANOVA can not.
```

3) Plot the lm with the data, with points colored by sex. (10 pts)


```{r}
ggplot(data = surveys_complete, mapping = aes(x = sex, y = weight, color = sex)) + geom_point()
```

4) Choose any model we've looked at so far, but use two predictor variables. Perform an LM, plot the results, and state if this changes your interpretation of the relationship between variables (10 pts)

```{r}
model_fit <- lm(hindfoot_length ~ sex, surveys_complete)
summary(model_fit)
```

```{r}
ggplot(data = surveys_complete, mapping = aes(x = sex, y = hindfoot_length, color = sex)) + geom_point()
```

```
No, it does not change my interpretation of the relationship between variables.
```

## Part Three


1) Add and commit this document (5 pts)

```
#Commands here
```

2) Push your changes to github (10 pts)

```
#Commands here
```



# MS students

My expectation is that you'll do this with your own data. If any part of this doesn't make sense with your data, please get in touch ASAP so we can work it out.
